---
title: "Learning mapping"
author: "ADS"
date: "`r `"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggmap)
library(ggplot2)
library(ggspatial)
library(ozmaps)
library(prettymapr)
library(readr)
library(rnaturalearth)
library(rosm)
library(sf)

```

Learning about mapping in R.

- ozmaps
- rosm
- rnaturalearth

More?

## Ozmaps

The simplest function call: `ozmap()`. This is a base R plot.

```{r}
ozmap()
```

But we can use the data in the package (itself extracted from ABS mapping data) with `ggplot()`.

```{r}
oz_states <- ozmaps::ozmap_states

oz_states

ggplot() +
  # this works without a mapping, because geom_sf() itself looks for a `geometry` column.
  geom_sf(data = oz_states) +
  coord_sf()

```

The package provides data for the following:

- `abs_ste`: states and territories
- `abs_ced`: Commonwealth electoral divisions
- `abs_lga`: local government areas

```{r}
ggplot(ozmaps::abs_lga) +
  geom_sf() +
  coord_sf()
```

Then zoom in using the affordances provided by `coord_sf()`

```{r}
ggplot(ozmaps::abs_lga) +
  geom_sf() +
  geom_sf_label(aes(label = NAME), size = 2) +
  coord_sf(
    # use the helper tool at Open Street Map
    xlim = c(150.1035, 151.5015), ylim = c(-33.3386, -34.1993)
  ) +
  theme_bw()
```

### Australian postcodes

```{r}
postcodes <- read_csv("aus_postcodes_geocoded.csv")

postcodes_shp <- read_sf(
  dsn = "1270055003_poa_2016_aust_shape/POA_2016_AUST.shp", 
  layer = "POA_2016_AUST"
)

ggplot() +
  geom_sf(data = abs_lga, fill = NA) +
  geom_sf(data = postcodes_shp, colour = "blue", fill = NA) +
  coord_sf(
    xlim = c(150.8, 151.5015), ylim = c(-33.55, -34.1)
  ) +
  geom_point(
    data = postcodes,
    aes(x = longitude, y = latitude),
    position = "jitter",
    colour = "red",
    alpha = 0.3,
    size = 1
  ) +
  theme_bw()


```

## ggmap

```{r}

sydney <- c( left = 150.8, bottom = -34.1, right = 151.5015, top= -33.55)

sydney_map <- get_stamenmap(sydney, zoom = 11, maptype = "terrain-background")

ggmap(sydney_map) +
  geom_point(
    data = postcodes,
    aes(x = longitude, y = latitude),
    position = "jitter",
    colour = "red",
    alpha = 0.4,
    size = 1
  )

```

Building with sf objects on top of a ggmap

```{r}

ggmap(sydney_map) +
  geom_sf(data = postcodes_shp, fill = NA, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(150.8, 151.5015), ylim = c(-33.55, -34.1)
  )

  
```

Terry Hills

```{r}

terrey_hills <- get_stamenmap(
  bbox = c(left = 151.1847, bottom = -33.7135, right = 151.2507, top = -33.6596),
  zoom = 13,
  maptype = "terrain"
)

ggmap(terrey_hills)

```



## rnaturalearth

```{r}
# Natural Earth shapefiles in the package rnaturalearth
world <- ne_countries(scale = "medium", returnclass = "sf")

ggplot(data = world) +
  # set up the world map with NE countries data
  geom_sf(fill = "antiquewhite") +
  # add labels for countries
  geom_sf_label(aes(label = name), size = 2) +
  # crop the map to the Mediterranean
  coord_sf(xlim = c(-1, 35), ylim = c(30, 47), expand = TRUE) +
  # adjust the overall visual appearance
  theme_bw() +
  theme(
    panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", size = 0.5),
    panel.background = element_rect(fill = "lightsteelblue1"),
    plot.caption = ggtext::element_textbox_simple(halign = 1.01, margin = ggplot2::margin(6, 0, 0, 0))
  ) +
  # add labels for the overall map
  labs(
    title = "Mediterranean map",
    subtitle = "Basemap for Mediterranean",
    caption = paste(
      "Basemap shapefiles from the Natural Earth project."
    )
  ) +
  # add direction and scale annotations from the package `ggspatial`
  annotation_scale(location = "bl") +
  annotation_north_arrow(
    location = "bl", 
    which_north = "true",
    pad_x = unit(1, "cm"),
    pad_y = unit(1, "cm")
  )

```



## rosm

For access to Bing Maps

```{r}

terrey_hills_bbox <- searchbbox("terrey hills, nsw")

prettymap({
  bmaps.plot(terrey_hills_bbox, type = "Aerial", key = NULL, zoomin = 0) # use Bing API key here
}, scale.label.col = "white")

```

